Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-06-11 emsoft-2019/V2/response.tex

\Mahmoud{\section{Response}	
We thank the reviewers for the detailed and useful feedback on our submission! We plan to improve our paper while taking into account the minor comments. Our response to your specific questions are summarized as follows.\\	
\textbf{Execution time for EMPCs(REVIEWER 4)}\\
The total time to evaluate the PWA function for a given $x_k$ in \autoref{eq:affine_map} consists of two steps:
\begin{itemize}
\item Selecting the region $\mathcal R_i$ such that $x_k\in \mathcal R_i$.
\item To evaluate the corresponding affine map $F_i x_k +G_i$. 
\end{itemize}
Evaluating the affine map is straightforward and requires less computations compared to the region selection step. Direct implementation of region selection (\autoref{lst:caseof}) is simple, but inefficient. More efficiency can be achieved by leveraging binary search tree structures (e.g. \cite{Mnnigmann:2011}). In general, explicit MPC is well-known for its high-speed implementation over embedded systems with low computational power. \cite{Johansen:2007} shows that explicit MPCs with hundreds of regions can be implemented on application specific integrated circuit (ASIC) with about $20000$ gates, leading to computation times in the order of $1\mu s$. In \cite{Bemporad:2006}, it is shown that for typical problems evaluating the explicit MPC takes significantly shorter time in comparison to solving on-line QP. \\
\textbf{Termination of algorithm for robust EMPC design (\autoref{lst:alg}) (REVIEWER 4)}\\
From theoretical point of view, the numerical error does not necessarily converge as each new designed controller may differ from the previous ones. Therefore, it is not easy to give an upper bound over the number of times that the robust MPC design needs to be repeated. However, one can use results on continuity of $argmin$ function in 
\begin{align*}
\label{eq:argmin}
&u_0(x_0,\Delta),\cdots,u_{N-1}(x_0,\Delta)=\\
&argmin \max_{w_0,\cdots,w_{N-1}} \sum_{i=0}^{N-1}(x_i^TQx_i+u_i^TRu_i) + x_N^TQ_Fx_N\nonumber\\
&\text{s.t.} \quad x_{i+1}=Ax_i+Bu_i + E w_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
%\quad &u_i=\mu x_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
&u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W}_{\Delta},\,\,\forall i\in\{0,1,\cdots,N\},
\end{align*}
 assuming its unique solution to every $\Delta$ and compute the Lipschitz constant for it over variations of $\Delta$. The computed Lipschitz constant will be only dependent on the dynamics of the system depicted in \autoref{eq:DSS} and hence can be used to evaluate the maximum error over incorrect region selection. From experimental point of view, using a reasonable $\varepsilon_{SAFE}$, our observations show that numerical error always has converged after only a few iterations.
}

### 2019-06-11 emsoft-2019/V2/Error_Analysis.tex
V2 folder is added+Rev 4 is answered

\begin{equation*}
\hat{\mathcal B}_{ij} := \left\{x\in\mathbb R^n\,|\, \| H_i x - K_i\| \le \varepsilon \right\}
\end{equation*}
where $\varepsilon$ captures the uncertainty in computing the correct region.
%
The width of the tube $\varepsilon$ is due to two sources of errors:
analog-to-digital conversion $\varepsilon_{A/D}$ and the quantization of region
bounds in memory $\varepsilon_{Q}$:
\begin{equation}\label{eq:epsilontot}
  \varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}

@@

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} = \|(H_i - \hat{H}_i) \hat{x} + (K_i - \hat{K}_i)\|
\end{equation}
where $\hat{x}$ is the finite-precision value of $x$.
%\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision ($H$ and $K$ matrices), and its counter-part quantized by $p$ bits 
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.

% Note that $\mathcal R_i\subset \bar{\mathcal R}_i$. In other words, $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$.
% Note that $\bar{\mathcal R}_i\cap \bar{\mathcal R}_j$ gives us the states where one of the two regions $\mathcal R_i, \mathcal R_i$ might be selected instead of the other one because of the fixed-precision implementation of the regions.   
% $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$ that include all states in $\mathcal R_i$this definition expands the regions $\mathcal $
 %That is, with infinite-precision arithmetic
 Therefore, the first term in~\autoref{eq:delta} can be computed less conservatively by maximizing only over 
 the expanded borders:
 \begin{equation}
 \label{eq:maximization}
 \max\{\|F_ix+G_i-F_jx-G_j\|,\,x\in \bar{\mathcal B}_{ij}\}
% \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|
 \end{equation}

@@

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} = |(H_i - \hat{H}_i) \qstatevar + (K_i - \hat{K}_i)|
\end{equation}
\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision (H and K), and its counter-part quantized in $p$ bits precision
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.
\end{comment}

@@

Once a quantized state $\hat x$ is given and a region $\mathcal R_i$ is chosen, computing $v_{i}$ itself introduces
imprecision, because, $v_{i}$ needs to be evaluated in finite-precision arithmetic:
\begin{equation}\label{eq:fperror}
  err(v_{i})_{p} = | (F_i - \hat{F}_{i}) \statevar + ( G_i - \hat{G}_{i} ) |
\end{equation}
where $\hat{F}$ and $\hat{G}$ represent the quantized values (in $p$ bits) for
the infinite precision values $F$ and $G$.

\autoref{eq:fperror} is evaluated for all \statevarmath that are inside the
region $\mathcal{R}_i$, together with all the values in the tube surrounding
region $\mathcal{R}_i$.
In this way we compute the error also for those points that belong to a neighbor
of $\mathcal{R}_i$, but because of finite-precision errors they are erroneously
mapped to control action $v_{i}$.

@@

\textcolor{red}{Sadegh: This is long comment on the content of this section. I cannot figure out how this line of reasoning is mapped to the implementation. We can discard my comment given that we are close to the deadline.}

\Sadegh{
I don't agree that we should add $\varepsilon_{A/D}$ and $\varepsilon_Q$ together. I understand that the implementation is done according to this, but I think we need to multiply $\varepsilon_{A/D}$ by the maximum norm of the coefficients $F_i$ in the controller. There are three approximations:
\begin{itemize}
\item using the quantized measured state $\hat x$ instead of the actual state $x$. We assume $\|x - \hat x\|\le \varepsilon_{A/D}$.
\item using $\hat H_i,\hat K_i$ which are the quantized versions of $H_i,K_i$ with $p$ bits precision. This is related to the boundaries of the regions.
\item using $\hat F_i,\hat G_i$ which are the quantized versions of $F_i,G_i$ with $p$ bits precision. This is related to the affine functions.
\end{itemize}
Let me define the sets $\mathcal H := \{H_i,K_i,i=1,2,\ldots, P\}$ and $\mathcal F:=\{F_i,G_i,i=1,2,\ldots, P\}$. Similarly, we define $\hat{\mathcal H}$ and $\hat{\mathcal F}$.
The controller (which is a piecewise affine function) can be written as the parametric function $\kappa(\mathcal H,\mathcal F,x)$. Our main goal is to make sure that
$$\|\kappa(\mathcal H,\mathcal F,x)-\kappa(\hat{\mathcal H},\hat{\mathcal F},\hat x)\|\le \Delta,\quad \forall x\in\mathcal X_{\mathsf s}.$$
We use tirangle these inequality:
\begin{align*}
\|\kappa(\mathcal H,\mathcal F,x)-\kappa(\hat{\mathcal H},\hat{\mathcal F},\hat x)
\|\le&\|\kappa(\mathcal H,\mathcal F,x)-\kappa(\hat{\mathcal H},\mathcal F,\hat x)\|\\
+&\|\kappa(\hat{\mathcal H},\mathcal F,\hat x)-\kappa(\hat{\mathcal H},\hat{\mathcal F},\hat x)\|.
\end{align*}
The second term is exactly what we have in section 4.2 (error in the evaluation of the affine function).
The first term can be further decomposed as
\begin{align*}
\|\kappa(\mathcal H,&\mathcal F,x)-\kappa(\hat{\mathcal H},\mathcal F,\hat x)\|\\
&\le \|\kappa(\mathcal H,\mathcal F,x)-\kappa(\mathcal H,\mathcal F,\hat x)\|+
 \|\kappa(\mathcal H,\mathcal F,\hat x) - \kappa(\hat{\mathcal H},\mathcal F,\hat x)\|\\
&\le \max_i\{\|F_i\|_2\}\|x-\hat x\| + \varepsilon_Q\\
& \le \max_i\{\|F_i\|_2\}\varepsilon_{A/D}+ \varepsilon_Q.
\end{align*}
The error $\varepsilon_{A/D}$ shows itself in the output of the controller by being amplified with the Lipschitz constant of the function which is $\max_i\{\|F_i\|_2\}$. 
As far as I can tell, $\varepsilon_{A/D}$ doesn't have anything to do with $\varepsilon_Q$. The computation of $\varepsilon_Q$ should be only with respect to the use of $\hat{\mathcal H}$ instead of $\mathcal H$ in the evaluation of constraints for the same value $\hat x$.
}
% The approximation error and the precision tuning for both borders and active
% functions are computed in Daisy. $\varepsilon_{Q}$ is given in input to the
% analysis. After we compute the incorrect region in MATLAB we obtain the value
% for max error $err(v_{i})$ from (8). Then we give the errors and the vectors F,G
% and H,K in input to Daisy. We used Daisy as a verification tool: we provide in
% input the expression we want to quantize (e.g. $\mathcal{R}_{i}$) together with
% the error bound we can tolerate (e.g. $\varepsilon_{Q}$), and Daisy returns a
% precision configuration (uniform or mixed) such that the input error is
% satisfied. The way it finds the right configuration for uniform precision is the
% following: it starts by checking whether the input expression, quantized in
% unary uniform precision (1bit), introduces an error bounded by the input error.
% If this is not the case, Daisy increases the precision format by 1 bit and check
% whether the error now is satisfied. This unary increment runs inside a loop that
% terminates when the first valid format is matched. This uniform word-length is
% going to be the baseline for the mixed-precision tuning phase. The
% mixed-precision algorithm keeps a list of variables to quantize, and collects
% different candidate configurations that satisfy the input error. It returns the
% best configuration based on the evaluation of a quality function.

>>>>

\begin{equation*}
\hat{\mathcal B}_{ij} := \left\{x\in\mathbb R^n\,|\, \| H_i x - K_i\| \le \varepsilon\,\, \text{ and }\,\, \| H_j x - K_j\| \le \varepsilon \right\},
\end{equation*}
where $\varepsilon$ captures the uncertainty in computing the correct region.
%
The width of the tube $\varepsilon$ is due to two sources of errors:
analog-to-digital conversion $\varepsilon_{A/D}$ and the quantization of region
bounds in memory $\varepsilon_{Q}$:
\begin{equation}\label{eq:epsilontot}
  \varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}

@@

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} \ge  \|(H_i - \hat{H}_i) \hat{x} + (K_i - \hat{K}_i)\|, \quad \forall \hat{x},\forall i,
\end{equation}
where $\hat{x}$ is the finite-precision value of $x$.
%\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision ($H$ and $K$ matrices), and its counter-part quantized by $p$ bits 
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.

% Note that $\mathcal R_i\subset \bar{\mathcal R}_i$. In other words, $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$.
% Note that $\bar{\mathcal R}_i\cap \bar{\mathcal R}_j$ gives us the states where one of the two regions $\mathcal R_i, \mathcal R_i$ might be selected instead of the other one because of the fixed-precision implementation of the regions.   
% $\bar{\mathcal R}_i$ is the expanded version of $\mathcal R_i$ that include all states in $\mathcal R_i$this definition expands the regions $\mathcal $
 %That is, with infinite-precision arithmetic
 Therefore, the first term in~\autoref{eq:delta} can be computed less conservatively by maximizing only over 
 the expanded borders:
 \begin{equation}
 \label{eq:maximization}
 \max_{x,i,j}\{\|F_ix+G_i-F_jx-G_j\|,\,x\in \bar{\mathcal B}_{ij}\}.
% \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|
 \end{equation}

@@

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} \ge  |(H_i - \hat{H}_i) \qstatevar + (K_i - \hat{K}_i)|, \forall  \qstatevar
\end{equation}
\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision (H and K), and its counter-part quantized in $p$ bits precision
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.
\end{comment}

@@

Once a quantized state $\hat x$ is given and a region $\mathcal R_i$ is chosen, computing $v_{i}$ itself introduces
imprecision, because, $v_{i}$ needs to be evaluated in finite-precision arithmetic:
\begin{equation}\label{eq:fperror}
  err(v_{i})_{p} \ge  | (F_i - \hat{F}_{i}) \statevar + ( G_i - \hat{G}_{i} ) |,\quad \forall \statevar\in \mathcal{R}_i\cup\bar{\mathcal B}_{ij},\,\,\forall j,
\end{equation}
where $\hat{F}$ and $\hat{G}$ represent the quantized values (in $p$ bits) for
the infinite precision values $F$ and $G$.

\autoref{eq:fperror} is evaluated for all \statevarmath that are inside the
region $\mathcal{R}_i$, together with all the values in the tube surrounding
region $\mathcal{R}_i$ (union of $\bar{\mathcal B}_{ij}$ for all $j$).
In this way we compute the error also for those points that belong to a neighbor
of $\mathcal{R}_i$, but because of finite-precision errors they are erroneously
mapped to control action $v_{i}$.

@@

\begin{comment}
\textcolor{red}{Sadegh: This is long comment on the content of this section. I cannot figure out how this line of reasoning is mapped to the implementation. We can discard my comment given that we are close to the deadline.}

\Sadegh{
I don't agree that we should add $\varepsilon_{A/D}$ and $\varepsilon_Q$ together. I understand that the implementation is done according to this, but I think we need to multiply $\varepsilon_{A/D}$ by the maximum norm of the coefficients $F_i$ in the controller. There are three approximations:
\begin{itemize}
\item using the quantized measured state $\hat x$ instead of the actual state $x$. We assume $\|x - \hat x\|\le \varepsilon_{A/D}$.
\item using $\hat H_i,\hat K_i$ which are the quantized versions of $H_i,K_i$ with $p$ bits precision. This is related to the boundaries of the regions.
\item using $\hat F_i,\hat G_i$ which are the quantized versions of $F_i,G_i$ with $p$ bits precision. This is related to the affine functions.
\end{itemize}
Let me define the sets $\mathcal H := \{H_i,K_i,i=1,2,\ldots, P\}$ and $\mathcal F:=\{F_i,G_i,i=1,2,\ldots, P\}$. Similarly, we define $\hat{\mathcal H}$ and $\hat{\mathcal F}$.
The controller (which is a piecewise affine function) can be written as the parametric function $\kappa(\mathcal H,\mathcal F,x)$. Our main goal is to make sure that
$$\|\kappa(\mathcal H,\mathcal F,x)-\kappa(\hat{\mathcal H},\hat{\mathcal F},\hat x)\|\le \Delta,\quad \forall x\in\mathcal X_{\mathsf s}.$$
We use tirangle these inequality:
\begin{align*}
\|\kappa(\mathcal H,\mathcal F,x)-\kappa(\hat{\mathcal H},\hat{\mathcal F},\hat x)
\|\le&\|\kappa(\mathcal H,\mathcal F,x)-\kappa(\hat{\mathcal H},\mathcal F,\hat x)\|\\
+&\|\kappa(\hat{\mathcal H},\mathcal F,\hat x)-\kappa(\hat{\mathcal H},\hat{\mathcal F},\hat x)\|.
\end{align*}
The second term is exactly what we have in section 4.2 (error in the evaluation of the affine function).
The first term can be further decomposed as
\begin{align*}
\|\kappa(\mathcal H,&\mathcal F,x)-\kappa(\hat{\mathcal H},\mathcal F,\hat x)\|\\
&\le \|\kappa(\mathcal H,\mathcal F,x)-\kappa(\mathcal H,\mathcal F,\hat x)\|+
 \|\kappa(\mathcal H,\mathcal F,\hat x) - \kappa(\hat{\mathcal H},\mathcal F,\hat x)\|\\
&\le \max_i\{\|F_i\|_2\}\|x-\hat x\| + \varepsilon_Q\\
& \le \max_i\{\|F_i\|_2\}\varepsilon_{A/D}+ \varepsilon_Q.
\end{align*}
The error $\varepsilon_{A/D}$ shows itself in the output of the controller by being amplified with the Lipschitz constant of the function which is $\max_i\{\|F_i\|_2\}$. 
As far as I can tell, $\varepsilon_{A/D}$ doesn't have anything to do with $\varepsilon_Q$. The computation of $\varepsilon_Q$ should be only with respect to the use of $\hat{\mathcal H}$ instead of $\mathcal H$ in the evaluation of constraints for the same value $\hat x$.
}
\end{comment}
% The approximation error and the precision tuning for both borders and active
% functions are computed in Daisy. $\varepsilon_{Q}$ is given in input to the
% analysis. After we compute the incorrect region in MATLAB we obtain the value
% for max error $err(v_{i})$ from (8). Then we give the errors and the vectors F,G
% and H,K in input to Daisy. We used Daisy as a verification tool: we provide in
% input the expression we want to quantize (e.g. $\mathcal{R}_{i}$) together with
% the error bound we can tolerate (e.g. $\varepsilon_{Q}$), and Daisy returns a
% precision configuration (uniform or mixed) such that the input error is
% satisfied. The way it finds the right configuration for uniform precision is the
% following: it starts by checking whether the input expression, quantized in
% unary uniform precision (1bit), introduces an error bounded by the input error.
% If this is not the case, Daisy increases the precision format by 1 bit and check
% whether the error now is satisfied. This unary increment runs inside a loop that
% terminates when the first valid format is matched. This uniform word-length is
% going to be the baseline for the mixed-precision tuning phase. The
% mixed-precision algorithm keeps a list of variables to quantize, and collects
% different candidate configurations that satisfy the input error. It returns the
% best configuration based on the evaluation of a quality function.

### 2019-06-11 emsoft-2019/V2/finite_precision.tex
V2 folder is added+Rev 4 is answered

In this work, we use the tool Daisy which implements roundoff error
analysis and mixed-precision tuning~\cite{DaisyTuning} for fixed-point arithmetic, and is
available as open-source. Fixed-point arithmetic implementations can choose different
rounding modes; here we consider truncation, which is more efficient than
rounding.

>>>>

In this work, we use the tool Daisy which implements roundoff error
analysis~\cite{Daisy} and mixed-precision tuning~\cite{DaisyTuning} for fixed-point arithmetic, and is
available as open-source. Fixed-point arithmetic implementations can choose different
rounding modes; here we consider truncation, which is more efficient than
rounding.
